Sensitivity Analysis

### 2024-05-24 Sections/Model.tex

			\subsubsection{Initial and boundary conditions} 
			It is assumed that the solvent is free of solute at the beginning of the process $c_{f0}=0$, that all the solid particles have the same initial solute content $c_{s0}$, and that the system is isothermal. The fluid at the inlet is also considered not to contain any solute. As the residence time is much shorter than the sampling time, the initial state estimate for the solute concentration in the fluid phase would be unreliable. The initial and boundary conditions are defined as follows:

@@

				{\footnotesize
					\begin{align*} \label{discretization}
						\dot{{\color{black}x}}(t) &= \cfrac{d x}{d t} = 
						\begin{bmatrix}
							\cfrac{d {\color{black}c_{f,1}}(t)}{d t} 	  \\
							\vdots					  \\
							\cfrac{d {\color{black}c_{f,N_z}}(t)}{d t} \\
							\\ \hline \\
							\cfrac{d {\color{black}c_{s,1}}(t)}{d t} 	  \\
							\vdots					  \\
							\cfrac{d {\color{black}c_{s,N_z}}(t)}{d t} \\
							\\ \hline \\
							\cfrac{d {\color{black}h_1}(t)}{d t} 	  \\
							\vdots 					  \\
							\cfrac{d {\color{black}h_{N_z}}(t)}{d t} \\
							\\ \hline \\
							\cfrac{d {\color{black}P}(t)}{d t} \\
							\\ \hline \\
							\cfrac{d {\color{black}y}(t)}{d t}
						\end{bmatrix}
						=
						\underbrace{\begin{bmatrix}
								{\color{black}G_1} \left( c_f,c_s,{\color{black}h}(t); \Theta \right)\\ 
								\vdots\\ 
								{\color{black}G_{N_z}} \left( c_f,c_s,{\color{black}h}(t); \Theta \right)\\ 
								\\ \hline \\ \\
								{\color{black}G_{N_z+1}} \left( c_f,c_s,{\color{black}h}(t); \Theta \right)\\ 
								\vdots\\
								{\color{black}G_{2N_z}} \left( {\color{black}c}(t),c_s,{\color{black}h}(t); \Theta \right)\\ 
								\\ \\ \hline \\ 
								{\color{black}G_{2N_z+1}} \left( c_f,c_s,{\color{black}h}(t); \Theta \right) \\
								\vdots\\
								{\color{black}G_{3N_z}} \left( c_f,c_s,{\color{black}h}(t); \Theta \right) \\ 
								\\ \\ \hline \\
								{\color{black}G_{3N_z+1}} \left( c_f,c_s,{\color{black}h}(t); \Theta \right) \\
								\\ \\ \hline \\
								{\color{black}G_{3N_z+2}} \left( c_f,c_s,{\color{black}h}(t); \Theta \right) 
						\end{bmatrix}}_{{\color{black}G} \left( x; \Theta \right)} 
				\end{align*} }
			where ${\color{black}x} \in \mathbb{R}^{N_x = 3N_z+2} $ and $\Theta \in \mathbb{R}^{N_\Theta =  N_{\theta} + N_u } $, $N_{\theta}$ is the number of parameters, $N_{u}$ is the number of control variables.

>>>>

			\subsubsection{Initial and boundary conditions} 
			It is assumed that the solvent is free of solute at the beginning of the process $c_{f0}=0$, that all the solid particles have the same initial solute content $c_{s0}$, and that the system is isothermal, hence the initial state is $h_0$. The fluid at the inlet is considered not to contain any solute. The initial and boundary conditions are defined as follows:

@@

				{\footnotesize
					\begin{align*} \label{discretization}
						\dot{x}(t) &= \cfrac{d x}{d t} = 
						\begin{bmatrix}
							\cfrac{d c_{f,1}(t)}{d t} 	  \\
							\vdots					  \\
							\cfrac{d c_{f,N_z}(t)}{d t} \\
							\\ \hline \\
							\cfrac{d c_{s,1}(t)}{d t} 	  \\
							\vdots					  \\
							\cfrac{d c_{s,N_z}(t)}{d t} \\
							\\ \hline \\
							\cfrac{d h_1(t)}{d t} 	  \\
							\vdots 					  \\
							\cfrac{d h_{N_z}(t)}{d t} \\
							\\ \hline \\
							\cfrac{d P(t)}{d t} \\
							\\ \hline \\
							\cfrac{d y(t)}{d t}
						\end{bmatrix}
						=
						\underbrace{\begin{bmatrix}
								G_1 \left( c_f,c_s,h; \Theta \right)\\ 
								\vdots\\ 
								G_{N_z} \left( c_f,c_s,h; \Theta \right)\\ 
								\\ \hline \\ \\
								G_{N_z+1} \left( c_f,c_s,h; \Theta \right)\\ 
								\vdots\\
								G_{2N_z} \left( c_f,c_s,h; \Theta \right)\\ 
								\\ \\ \hline \\ 
								G_{2N_z+1} \left( c_f,c_s,h; \Theta \right) \\
								\vdots\\
								G_{3N_z} \left( c_f,c_s,h; \Theta \right) \\ 
								\\ \\ \hline \\
								G_{3N_z+1} \left( c_f,c_s,h; \Theta \right) \\
								\\ \\ \hline \\
								G_{3N_z+2} \left( c_f,c_s,h; \Theta \right) 
						\end{bmatrix}}_{G \left( x; \Theta \right)} 
				\end{align*} }
			where ${\color{black}x} \in \mathbb{R}^{N_x = 3N_z+2} $ and $\Theta \in \mathbb{R}^{N_\Theta =  N_{\theta} + N_u } $, $N_{\theta}$ is the number of parameters, $N_{u}$ is the number of control variables.

### 2024-05-24 Sections/Sensitivity_Analysis.tex
Updates from Overleaf

	Local derivative-based methods involve taking the total derivative of the state vector $x$ with respect to parameters $p$, where $p$ is a subset of the parameter space $\theta$. This set of derivatives, known as sensitivity equations, is integrated simultaneously with the process model. The sensitivity analysis investigates how responsive the solution is for the perturbation of the parameter $p$. As discussed by \citet{Dickinson1976}, the sensitivity analysis can be used to determine the influence of the uncertainty on the solution of the original system. A sensitivity analysis can be used to distinguish sensitive parameters from insensitive ones, which might be helpful for model reduction. Finally, from a control engineering point of view, the sensitivity analysis allows sorting the control variables with respect to the level of effort required to change the model's output.

@@

	{\footnotesize
		\begin{align}
			\label{SA_def}
			{Z}(x(t);p) &= \cfrac{dx(t)}{dp}
	\end{align} }

@@

	The sensitivity equation can be obtained by applying the definition of the total derivative to the Equation \ref{SA_dt}.

@@

	The sensitivity Equation \ref{SA_eq_full} is solved simultaneously with the original system and is made of three terms: Jacobian ${J_x}(x(t);p)$, sensitivity matrix $S(x(t);p)$ and Jacobian ${J_p}(x(t);p)$. The Jacobian ${J_x}(x(t);p)$ represents the matrix of equations of size $N_x \times N_x$, where each equation ${J_x}(n_x,n_x)$ is the derivative of process model equations ${G}_{n_x}(x(t);p)$ with respect to the state variable $x_{n_p}$.

@@

	The sensitivity matrix $S(x(t);p)$ represents the matrix of equations of size $N_x \times N_p$, where each subequation $S(n_x,n_p)$ is the derivative of the state variable $x_{n_x}$ with respect to the parameter $p_{n_p}$.

@@

	The Jacobian ${J_p}(x(t);p)$ represents the matrix of equations of size $N_x \times N_p$, where each subequation ${J_p}(n_x,n_p)$ is the partial derivative of the process model equation $F_{n_x}$ with respect to the parameter $p_{n_p}$.

@@

	{\footnotesize
		\begin{align}
			{\textbf{G}}\left(x(t_0);p\right)  = 
			\begin{bmatrix}x(t_0),						& 
				\cfrac{ \text{d}x(t_0) }{ d{p_1} },		& 
				\cdots,					 				&
				\cfrac{ dx(t_0) }{ d{p_{N_p}} } 
			\end{bmatrix}^T
	\end{align} }

>>>>

	Local derivative-based methods involve taking the total derivative of the state vector $x$ with respect to parameters $p$, where $p$ is a subset of the parameter space $\theta$. The set of derivatives, known as sensitivity equations, is integrated simultaneously with the process model. The sensitivity analysis shows how responsive the solution is for changes in the parameter $p$. As discussed by \citet{Dickinson1976}, the sensitivity analysis can be used to determine the influence of the uncertainty on the solution of the original system. A sensitivity analysis can be used to distinguish sensitive parameters from insensitive ones, which might be helpful for model reduction. Finally, from a control engineering point of view, the sensitivity analysis allows sorting the control variables with respect to the level of effort required to change the model's output.

@@

	{\footnotesize
		\begin{equation}
			\label{SA_def}
			Z(x(t);p) = \cfrac{dx(t)}{dp}
	   \end{equation} }

@@

	The sensitivity equation can be obtained by applying the definition of the total derivative to Equation \ref{SA_dt}.

@@

	Equation \ref{SA_eq_full} is solved simultaneously with the process model and consists of three terms: Jacobian ${J_x}(x(t);p)$, sensitivity matrix $S(x(t);p)$ and Jacobian ${J_p}(x(t);p)$. The Jacobian ${J_x}(x(t);p)$ represents the matrix of equations of size $N_x \times N_x$, where each equation ${J_x}(n_x,n_x)$ is the derivative of ${G}_{n_x}(x(t);p)$ with respect to the state variable $x_{n_p}$.

@@

	The sensitivity matrix $S(x(t);p)$ represents the matrix of equations of size $N_x \times N_p$, where each subequation $S(n_x,n_p)$ is the derivative of the state variable $x_{n_x}$ with respect to the parameter $p_{n_p}$. Matrix ${J_x}(x(t);p)$ and $S(x(t);p)$ describe indirect influence of $p_{n_p}$ on the model output.

@@

	The Jacobian ${J_p}(x(t);p)$ represents the matrix of equations of size $N_x \times N_p$, where each subequation ${J_p}(n_x,n_p)$ is the partial derivative of the process model equation $F_{n_x}$ with respect to the parameter $p_{n_p}$. ${J_p}(n_x,n_p)$ can be defines direct effect of $p_{n_p}$ on the model output.

@@

	{\footnotesize
		\begin{align}
			{\textbf{G}}\left(x(t_0);p\right)  &= 
			\begin{bmatrix}
                x(t_0),						               &
				\cfrac{ \text{d}x(t_0) }{ d{p_1} },		   &
				\cdots,					 				   &
				\cfrac{ dx(t_0) }{ d{p_{N_p}} }            
			\end{bmatrix}^\top = \\ &=
            \begin{bmatrix} 
                \quad x_0,	                               &
				\quad ~0,		                           &
			    \quad \cdots,			                   &
				\quad 0 \quad~~
			\end{bmatrix}^\top
	\end{align} }
